\vspace{-2.5em} Relating Spherical Coordinates In Rotated Frames\vspace{-1em}

### 2017-06-27 notes/2017-06-09-rotated-frames/report/report.tex
Fixed incorrect reference.
Correct error in spherical coordinates equations. Fix phi cases.

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Relating Spherical Coordinates In Rotated Frames\vspace{-1em}} \author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}June 9, 2017\vspace{-1em}}
\maketitle

@@

\section{Relating Spherical Coordinates In Two Oblique Frames}
First, we expand a single unit vector $\mh{r}$ in two different coordinate systems
\begin{align}
  \mh{r} = x\mh{x} + y\mh{y} + z\mh{z} &= x'\mh{x}' + y'\mh{y}' + z'\mh{z}'.
\intertext{Next, we rewrite the vector in spherical coordinates}
  \mh{r} = \cos\phi\sin\theta\mh{x} + \sin\phi\sin\theta\mh{y} + \cos\theta\mh{z} &= \cos\phi'\sin\theta'\mh{x}' + \sin\phi'\sin\theta'\mh{y}' + \cos\theta'\mh{z}'.\label{eq:sphere}
\end{align}
Finally, we relate the coordinate systems. We will only consider coordinate
systems that are related by a right handed rotation by angle $\psi$ about the
$\mh{y}$ axis, so the coordinate systems are related by
\begin{subequations}
\begin{align}
  \mh{x} &= \cos\psi\mh{x}' +\sin\psi\mh{z}'\\
  \mh{y} &= \mh{y}'\\
  \mh{z} &= -\sin\psi\mh{x}' + \cos\psi\mh{z}'.
 \end{align}\label{eq:coords}
\end{subequations}%
Plugging equation \ref{eq:coords} into equation \ref{eq:sphere} and equating
each component gives
\begin{subequations}
  \begin{align}
    \cos\psi\cos\phi\sin\phi - \sin\psi\cos\theta\label{eq:preva} &=\cos\phi'\sin\theta'\\
     \sin\phi\sin\theta&=\sin\phi'\sin\theta' \label{eq:prevb}\\
    \sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta&=\cos\theta' .\label{eq:prevc}
  \end{align}
\end{subequations}
Solving for $\theta'$ and $\phi'$ in terms of $\theta$, $\phi$, and $\psi$ gives
\begin{subequations}
\begin{align}
  \theta' &= \arccos\left(\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta\right)\label{eq:thetap}\\
  \phi' &= \arccos\left(\frac{\cos\psi\cos\phi\sin\phi - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right).\label{eq:phip}
\end{align}\label{eq:solution}% 
\end{subequations}
Equation \ref{eq:thetap} follows directly from equation \ref{eq:prevc}. Equation
\ref{eq:phip} follows from plugging equation \ref{eq:thetap} into equation \ref{eq:prevc} and using the identity $\sin(\arccos(x)) = \sqrt{1 - x^2}$.

The inverse equations can be found by the substituting $\psi \rightarrow -\psi$:
\begin{subequations}
\begin{align}
  \theta &= \arccos\left(-\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta'\right)\\
  \phi &= \arccos\left(\frac{\cos\psi\cos\phi'\sin\phi' + \sin\psi\cos\theta'}{\sqrt{1 - (\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta')^2}}\right).
\end{align}\label{eq:solutionp}% 
\end{subequations}

>>>>

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Relating Spherical Coordinates In Rotated Frames\vspace{-1em}} \author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}June 9, 2017\\ (Updated: June 27, 2017)\vspace{-1em}}
\maketitle

@@

\section{Relating Spherical Coordinates In Two Oblique Frames}
First, we expand a single unit vector $\mh{r}$ in two different coordinate systems
\begin{align}
  \mh{r} = x\mh{x} + y\mh{y} + z\mh{z} &= x'\mh{x}' + y'\mh{y}' + z'\mh{z}'.
\intertext{Next, we rewrite the vector in spherical coordinates}
  \mh{r} = \cos\phi\sin\theta\mh{x} + \sin\phi\sin\theta\mh{y} + \cos\theta\mh{z} &= \cos\phi'\sin\theta'\mh{x}' + \sin\phi'\sin\theta'\mh{y}' + \cos\theta'\mh{z}'.\label{eq:sphere}
\end{align}
Finally, we relate the coordinate systems. We will only consider coordinate
systems that are related by a right handed rotation by angle $\psi$ about the
$\mh{y}$ axis, so the coordinate systems are related by
\begin{subequations}
\begin{align}
  \mh{x} &= \cos\psi\mh{x}' +\sin\psi\mh{z}'\\
  \mh{y} &= \mh{y}'\\
  \mh{z} &= -\sin\psi\mh{x}' + \cos\psi\mh{z}'.
 \end{align}\label{eq:coords}
\end{subequations}%
Plugging equation \ref{eq:coords} into equation \ref{eq:sphere} and equating
each component gives
\begin{subequations}
  \begin{align}
    \cos\psi\cos\phi\sin\phi - \sin\psi\cos\theta\label{eq:preva} &=\cos\phi'\sin\theta'\\
     \sin\phi\sin\theta&=\sin\phi'\sin\theta' \label{eq:prevb}\\
    \sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta&=\cos\theta' .\label{eq:prevc}
  \end{align}
\end{subequations}
Solving for $\theta'$ and $\phi'$ in terms of $\theta$, $\phi$, and $\psi$ gives
\begin{subequations}
\begin{align}
  \theta' &= \arccos\left(\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta\right)\label{eq:thetap}\\
  \phi' &= \arccos\left(\frac{\cos\psi\cos\phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right).\label{eq:phip}
\end{align}\label{eq:solution}% 
\end{subequations}
Equation \ref{eq:thetap} follows directly from equation \ref{eq:prevc}. Equation
\ref{eq:phip} follows from plugging equation \ref{eq:thetap} into equation \ref{eq:preva} and using the identity $\sin(\arccos(x)) = \sqrt{1 - x^2}$.

The inverse equations can be found by the substituting $\psi \rightarrow -\psi$:
\begin{subequations}
\begin{align}
  \theta &= \arccos\left(-\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta'\right)\\
  \phi &= \arccos\left(\frac{\cos\psi\cos\phi'\sin\theta' + \sin\psi\cos\theta'}{\sqrt{1 - (\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta')^2}}\right).
\end{align}\label{eq:solutionp}% 
\end{subequations}

\textbf{Note:} The arccos in equations \ref{eq:solution} and \ref{eq:solutionp}
depends the range of $\phi$. If $\phi \in [-\pi, \pi]$ then use
\begin{align}
  \phi' &=
          \begin{cases}
            \arccos\left(\frac{\cos\psi\cos\phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right) \ \ \ &0 \leq \phi < \pi  \\
            -\arccos\left(\frac{\cos\psi\cos\phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right) \ \ \ &-\pi \leq \phi < 0\\
          \end{cases}
\end{align}
